Flabby and injective objects in toposes

### 2018-06-19 paper-flabby-objects.tex
typo
Lots of minor improvements

\newcommand{\xra}[1]{\xrightarrow{#1}}
\newcommand{\XXX}[1]{\textbf{XXX: #1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\op}{\mathrm{op}}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Mod}{Mod}
\DeclareMathOperator{\Sh}{Sh}
\DeclareMathOperator{\PSh}{PSh}
\newcommand{\Set}{\mathrm{Set}}
\renewcommand{\_}{\mathpunct{.}\,}

@@

If~$F$ is a flabby sheaf in the traditional sense, then~$F$ is obviously also
flabby in the sense of Definition~\ref{defn:flabby-sheaf} -- singleton
coverings will do. Conversely, let~$F$ be a flabby sheaf in the sense of
Definition~\ref{defn:flabby-sheaf}. Let~$s \in F(U)$ be a local section. Zorn's
lemma implies that there is a maximal extension~$s' \in F(U')$. By assumption,
there is an open covering~$X = \bigcup_i U_i$ such that, for all~$i$, the
section~$s'$ can be extended to~$U' \cup U_i$. Since~$s'$ is maximal, $U' \cup
U_i = U'$ for all~$i$. Therefore~$X = \bigcup_i U_i \subseteq U'$; hence~$s'$ is a
global section, as desired.

@@

\section{Flabby sets}

We intend this section to be applied in the internal language of an elementary
topos. We will therefore reason purely intuitionistically. We
adopt the terminology of~\cite{kock:partial-maps} regarding subterminals and
subsingletons: A subset~$K \subseteq X$ is \emph{subterminal} if and only if any given
elements are equal ($\forall x,y \in K\_ x = y$), and it is a
\emph{subsingleton} if and only if there is an element~$x \in X$ such that~$K
\subseteq \{ x \}$. Any subsingleton is trivially subterminal, but the converse
might fail.

@@

In the presence of the law of excluded middle, a set is flabby if and only if
it is inhabited. In fact, this characterization is a constructive taboo:

\begin{prop}If any inhabited set is flabby, then the law of excluded middle
holds.
\end{prop}

@@

\begin{defn}A set~$X$ is \emph{injective} if and only if, for any injection~$i
: A \to B$, any map~$f : A \to X$ can be extended to a map~$B \to X$.
An~$R$-module~$M$ is \emph{injective} if and only if, for any linear
injection~$i : A \to B$ between~$R$-modules, any linear map~$f : A \to M$ can
be extended to a linear map~$B \to X$.\end{defn}

@@

\begin{proof}\begin{enumerate}
\item One can check that, for instance, the full powerset~$\P(X)$ and the set of
subterminals~$\P_{\leq 1}(X)$ are each injective.~\cite{XXX}
\item The forgetful functor from modules to sets possesses a left exact left
adjoint. More explicitly, if~$i : A \to B$ is an injective map between sets and
if~$f : A \to M$ is an arbitrary map, then the induced map~$R\langle A \rangle
\to R\langle B \rangle$ between free modules is also injective, the given
map~$f$ lifts to a linear map~$R\langle A \rangle \to M$, and an
extension~$R\langle B \rangle \to M$ induces an extension~$B \to M$ of~$f$.
\item One verifies that any abelian group can be embedded into a divisible
abelian group. By the axiom of choice, divisible abelian groups are injective.
The result for modules over arbitrary rings then follows purely
formally~\cite{XXX}. \qedhere
\end{enumerate}\end{proof}

@@

@@

\begin{proof}We first cover the case of sets. Let~$K \subseteq I^T$ be a
subterminal. We consider the injectivity diagram
\[ \xymatrix{
  T' \ar[r]\ar[d] & T \ar@{-->}[ld] \\
  I
} \]
where~$T'$ is the subset~$\{ s \in T \,|\, \text{$K$ is inhabited} \} \subseteq T$ and the
solid vertical map sends~$s \in T'$ to~$g(s)$, where~$g$ is an arbitrary element
of~$K$. This association is well-defined. Since~$I$ is injective, a dotted lift
as indicated exists. If~$K$ is inhabited, this lift is an element of~$K$.

The same kind of argument applies to the case of modules. If~$K$ is a
subterminal of~$\Hom_R(T,I)$, we define~$T'$ to be the submodule
$\{ s \in I \,|\, \text{$s = 0$ or $K$ is inhabited} \}$ and consider the
analogous injectivity diagram, where the solid vertical map~$f : T' \to I$ is now
defined by cases: Let~$s \in T'$. If~$s = 0$, then we set~$f(s) = 0$; if~$K$ is
inhabited, then we set~$f(s) \defeq g(s)$, where~$g$ is an arbitrary element
of~$K$. This association is again well-defined, and a dotted lift yields the
desired element of~$\Hom_R(T,I)$.
\end{proof}

@@

\begin{proof}Set~$X \defeq \{ \bar{f} \in I^B \,|\, \bar{f} \circ i =
f \}$. Let~$K \subseteq X$ be a subsingleton. We consider the injectivity diagram
\[ \xymatrix{
  i[A] \cup B' \ar@{^{(}->}[r]\ar[d]_g & B \ar@{-->}[ld] \\
  I
} \]
where~$B'$ is the set~$\{ s \in B \,|\, \text{$K$ is inhabited} \}$ and the solid
vertical arrow~$g$ is defined in the following way: Let~$s \in i[A] \cup B'$.
If~$s \in i[A]$, then we set~$g(s) \defeq f(a)$, where~$a \in A$ is an element such
that~$s = i(a)$. If~$s \in B'$, then we set~$g(s) \defeq \bar{f}(s)$,
where~$\bar{f}$ is any element of~$K$. These prescriptions determine a well-defined
map.

@@

\begin{proof}The proof is similar to the proof of
Proposition~\ref{prop:set-of-extensions-flabby}. We
set~$X \defeq \{ \bar{f} \in \Hom_R(B,I) \,|\, \bar{f} \circ i =
f \}$. Let~$K \subseteq X$ be a subsingleton. We consider the injectivity diagram
\[ \xymatrix{
  i[A] + B' \ar@{^{(}->}[r]\ar[d]_g & B \ar@{-->}[ld] \\
  I
} \]
where~$B'$ is the submodule~$\{ s \in B \,|\, \text{$s = 0$ or $K$ is
inhabited} \} \subseteq B$ and the solid vertical arrow~$g$ is defined in the following
way: Let~$s \in i[A] + B'$. Then~$s = i(a) + t$ for an element~$a \in A$ and an
element~$t \in B'$. Since~$t \in B'$, $t = 0$ or~$K$ is inhabited. If~$t = 0$,
we set~$g(s) \defeq f(a)$. If~$K$ is inhabited, we set~$g(s) \defeq f(a) +
\bar{f}(s)$, where~$\bar{f}$ is any element of~$K$. These prescriptions
determine a well-defined map.

@@

\begin{proof}For the first statement, let~$X \defeq \{ u \in M \,|\, p(u) = s
\}$. Let~$K \subseteq X$ be a subterminal. Since~$p$ is surjective, there is a
preimage~$u_0 \in X$. The translated set~$K - u_0 \subseteq M$ is still a
subterminal, and its preimage under~$i$ is as well. Since~$M'$ is flabby, there
is an element~$v \in M'$ such that~$i^{-1}[K - u_0] \subseteq \{v\}$. We verify
that~$K \subseteq \{u_0 + i(v)\}$.

@@

@@

Instead of referencing arbitrary stages~$A \in \E$, one can also just reference
the generic stage: Let~$\P_{\leq1}(X)$ denote the \emph{object of subterminals}
of~$X$; this object is a certain suboject of~$\P(X) = [X,\Omega_\E]_\E$, the
powerobject of~$X$. The subobject~$K_0$ of~$\P_{\leq1}(X) \times X$ classified by the
evaluation morphism~$\P_{\leq1}(X) \times X \to \P(X) \times X \to \Omega_\E$
is the \emph{generic subterminal} of~$X$. The object~$X$ is flabby if and only
if there exists an epimorphism~$B \to \P_{\leq1}(X)$ and a morphism~$B \to X$
such that the following diagram commutes.
\[ \xymatrix{
  K_0 \times_{\P_{\leq1}(X)} B \ar@{^{(}->}[r]\ar[d] & \P_{\leq1}(X) \ar@{-->}@/^/[ld] \\
  X
} \]

@@

\begin{proof}These properties are shared by any property which can be defined
in the stack semantics~\cite[Lemma~7.3]{shulman:stack-semantics}.
\end{proof}

@@

@@

\begin{proof}Let~$X \in \F$ be a flabby object. Without loss of generality, it
suffices to verify condition~$({\star})$ for the terminal stage~$A = 1 \in \E$.
Let~$k : K \to 1$ be a monomorphism in~$\E$ and let~$x : K \to f_*(X)$ be an arbitrary morphism.
Then~$f^*(k) : f^*(K) \to 1$ is a monomorphism in~$\F$ and~$x^t : f^*(K) \to X$
is a morphism in~$\F$. Since~$X$ is flabby, there is an epimorphism~$B \to 1$
in~$\F$ and a morphism~$y : B \to X$ such that the morphism~$f^*(K) \times B
\to X$ factors over~$y$. Hence~$x$ factors over~$f_*(y) : f_*(B) \to f_*(X)$.
By assumption, the morphism~$f_*(B) \to f_*(1)$ is an epimorphism.
\end{proof}

The assumption on~$f_*$ of Proposition~\ref{prop:pushforward-of-flabby-objects}
is for instance satisfied if~$f$ is a local geometric morphism or if~$f$ is
induced by an open continuous injection between topological spaces. % XXX

@@

\begin{prop}\label{prop:notions-of-internal-injectivity}
Let~$\E$ be an elementary topos. Then the following statements about an
object~$I \in \E$ are equivalent.
\begin{enumerate}
\item $I$ is internally injective.
\item The functor~$[\cdot, I] : \E^\op \to \E$ maps monomorphisms in $\E$
to morphisms for which any global element of the target locally (after change of
base along an epimorphism) possesses a preimage.
\item For any morphism $p : A \to 1$ in $\E$, the object $p^*(I)$ has property~(1)
as an object of $\E/A$.
\item For any morphism $p : A \to 1$ in $\E$, the object $p^*(I)$ has property~(2)
as an object of $\E/A$.
\item From the point of view of the internal language of~$\E$, the object~$I$
is an injective set.
\end{enumerate}
\end{prop}

\begin{proof}
The implications ``(1)~$\Rightarrow$~(2)'', ``(3)~$\Rightarrow$~(4)'',
``(3)~$\Rightarrow$~(1)'', and ``(4)~$\Rightarrow$~(2)'' are trivial.

The equivalence ``(3)~$\Leftrightarrow$~(5)'' follows directly from the
interpretation rules of the stack semantics.

The implication ``(2)~$\Rightarrow$~(4)'' employs the
extra left adjoint $p_! : \E/A \to \E$ of $p^* : \E
\to \E/A$~(which maps an object~$(X \to A)$ to~$X$), as in the usual proof that
injective sheaves remain injective when
restricted to smaller open subsets: We have that $p_* \circ [\cdot, p^*(I)]_{\E/A}
\cong [\cdot, I]_\E \circ p_!$, the functor $p_!$ preserves monomorphisms, and one
can check that $p_*$ reflects the property that global elements locally possess
preimages. Details are in~\cite[Thm.~1.1]{harting}.\footnote{Harting formulates
the statement for abelian group objects, and has to assume that~$\E$ contains a
natural numbers object to ensure the existence of an abelian version of~$p_!$.}

The implication ``(4)~$\Rightarrow$~(3)'' follows by performing an extra change of
base, since any non-global element becomes a global element after a suitable
change of base.
\end{proof}

Let~$R$ be a ring in~$\E$. Then the analogue of
Proposition~\ref{prop:notions-of-internal-injectivity} holds for~$R$-modules
in~$\E$, if~$\E$ is assumed to have a natural numbers object. The extra
assumption is needed in order to construct the left adjoint~$p_! :
\Mod_{\E/A}(R \times A) \to \Mod_\E(R)$. Phrased in the internal language, this
adjoint maps a family~$(M_a)_{a \in A}$ of~$R$-modules to the direct
sum~$\bigoplus_{a \in A} M_a$. Details on this construction, phrased in the
language of sets but interpretable in the internal language, can for instance
be found in~\cite[page~54]{mines-richman-ruitenburg:constructive-algebra}.
It seems that Harting was not aware that this construction is possible, as she
(mistakingly) states in~\cite[page~233]{harting:remark} that ``the notions of
injectivity and internal injectivity do not coincide'' for modules.

@@

For the ``if'' direction, write~$\E = \Sh(X)$ and let~$I$ be an internally
injective object. Let~$i : A \to B$ be a monomorphism in~$\Sh(X)$ and let~$f :
A \to I$ be an arbitrary morphism. We want to show that there exists an
extension $B \to I$ of~$f$ along~$i$. To this end, we consider the sheaf of
such extensions, defined by the internal expression
\[ F \defeq \{ \bar{f} \in [B,I] \,|\, \bar{f} \circ i = f \}. \]
Global elements of~$F$ are extensions of the kind we are looking for.
By Proposition~\ref{prop:set-of-extensions-flabby}, this sheaf is a flabby
object and therefore by Proposition~\ref{prop:flabby-sheaves-objects} a flabby
sheaf. Zorn's lemma implies that~$F$ has a global element.
\end{proof}

@@

\begin{scholium}For any elementary topos~$\E$:
\begin{enumerate}
\item Any object can be embedded into an internally injective object.
\item The underlying unstructured object of an internally injective module is
internally injective.
\item Any internally injective object is flabby.
\item Any object can be embedded into a flabby object.
\item The terminal object is flabby. The product of flabby objects if flabby.
\item Let~$I$ be an internally injective object. Let~$T$ be an arbitrary
object. Then~$[T,I]$ is a flabby object.
\item (If~$\E$ has a natural numbers object.) Let~$I$ be an internally
injective~$R$-module. Let~$T$ be an arbitrary~$R$-module. Then~$[T,I]_R$, the
subobject of the internal Hom consisting only of the linear maps, is a flabby
object.
\item Let~$0 \to M' \to M \to M'' \to 0$ be a short exact sequence
of~$R$-modules in~$\E$. If~$M'$ and~$M''$ are flabby objects, so is~$M$.
\end{enumerate}
\end{scholium}

@@

\begin{proof}We only have to verify exactness at~$\Gamma(M'')$, so let~$s \in
\Gamma(M')$. Interpreting Proposition~\ref{prop:set-of-preimages-flabby}
in~$\E$, we see that the object of preimages of~$s$ is object. Since~$\E$ is
localic, this object is a flabby sheaf; since Zorn's lemma is available, it
possesses a global element. Such an element is the desired preimage of~$s$
in~$\Gamma(M)$.\end{proof}

@@

\begin{rem}A generalization of the traditional notion of a flabby sheaf to
elementary toposes is the following. An object~$X$ of an elementary topos~$\E$
is \emph{strongly flabby} if and only if, for any monomorphism~$K \to 1$
in~$\E$, any morphism~$K \to X$ lifts to a morphism~$1 \to X$.

@@

% EXT, TOR
% Higher direct images

@@

The main obstacle preventing Grothendieck's theory to be interpreted
constructively is its reliance on injective resolutions. It is known that in
the absence of the axiom of choice, much less in a purely intuitionistic
context, there might not be any nontrivial injective abelian
group~\cite{blass:inj-proj-axc}.

In principle, this problem could be remedied by employing flabby resolutions
instead of injective ones. There are, however, two problems with this
suggestion. Firstly, we needed Zorn's lemma to show that flabby sheaves are
acyclic for the global sections functor (Proposition~XXX); however, we noticed
the substitute property in XXX which might be good enough for further
developments. The more serious problem is that it is an open question whether
one can show, purely intuitionistically, that any sheaf of modules embeds into
a flabby sheaf of modules. The following is known about this problem:

>>>>

\newcommand{\xra}[1]{\xrightarrow{#1}}
\newcommand{\XXX}[1]{\textbf{XXX: #1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\op}{\mathrm{op}}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Mod}{Mod}
\DeclareMathOperator{\Sh}{Sh}
\DeclareMathOperator{\PSh}{PSh}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\renewcommand{\_}{\mathpunct{.}\,}

@@

% Take care: This paragraph is referenced in prop:global-elements.
If~$F$ is a flabby sheaf in the traditional sense, then~$F$ is obviously also
flabby in the sense of Definition~\ref{defn:flabby-sheaf} -- singleton
coverings will do. Conversely, let~$F$ be a flabby sheaf in the sense of
Definition~\ref{defn:flabby-sheaf}. Let~$s \in F(U)$ be a local section. Zorn's
lemma implies that there is a maximal extension~$s' \in F(U')$. By assumption,
there is an open covering~$X = \bigcup_i U_i$ such that, for all~$i$, the
section~$s'$ can be extended to~$U' \cup U_i$. Since~$s'$ is maximal, $U' \cup
U_i = U'$ for all~$i$. Therefore~$X = \bigcup_i U_i \subseteq U'$; hence~$s'$ is a
global section, as desired.

@@

\section{Flabby sets}\label{sect:flabby-sets}

We intend this section to be applied in the internal language of an elementary
topos; we will speak about sets and maps between sets, but intend our arguments
to be applied to objects and morphisms in toposes. We will therefore be careful
to reason purely intuitionistically. We
adopt the terminology of~\cite{kock:partial-maps} regarding subterminals and
subsingletons: A subset~$K \subseteq X$ is \emph{subterminal} if and only if any given
elements are equal ($\forall x,y \in K\_ x = y$), and it is a
\emph{subsingleton} if and only if there is an element~$x \in X$ such that~$K
\subseteq \{ x \}$. Any subsingleton is trivially subterminal, but the converse
might fail.

@@

In the presence of the law of excluded middle, a set is flabby if and only if
it is inhabited. This characterization is a constructive taboo:

\begin{prop}\label{prop:taboo}
If any inhabited set is flabby, then the law of excluded middle
holds.
\end{prop}

@@

\begin{defn}\begin{enumerate}
\item A set~$X$ is \emph{injective} if and only if, for any injection~$i
: A \to B$, any map~$f : A \to X$ can be extended to a map~$B \to X$.
\item An~$R$-module~$M$ is \emph{injective} if and only if, for any linear
injection~$i : A \to B$ between~$R$-modules, any linear map~$f : A \to M$ can
be extended to a linear map~$B \to M$.
\end{enumerate}
\[ \xymatrix{
  A \ar@{^{(}->}[r]\ar[d] & B \ar@{-->}@/^/[ld] &&
  A \ar@{^{(}->}[r]\ar[d] & B \ar@{-->}@/^/[ld] \\
  X &&& M
} \]
\end{defn}

@@

\begin{proof}\begin{enumerate}
\item One can check that, for instance, the full powerset~$\P(X)$ and the set of
subterminals~$\P_{\leq 1}(X)$ are each injective.~\cite{XXX}
\item The forgetful functor from modules to sets possesses a left exact left
adjoint. More explicitly, if~$i : A \to B$ is an injective map between sets and
if~$f : A \to M$ is an arbitrary map, then the induced map~$R\langle A \rangle
\to R\langle B \rangle$ between free modules is also injective, the given
map~$f$ lifts to a linear map~$R\langle A \rangle \to M$, and an~$R$-linear
extension~$R\langle B \rangle \to M$ induces an extension~$B \to M$ of~$f$.
\item One verifies that any abelian group can be embedded into a divisible
abelian group. By the axiom of choice, divisible abelian groups are injective.
The result for modules over arbitrary rings then follows purely
formally~\cite{XXX}. \qedhere
\end{enumerate}\end{proof}

@@

A further corollary of Proposition~\ref{prop:injective-flabby} is that the
statement ``any inhabited set is injective'' is a constructive taboo: If any
inhabited set is injective, then any inhabited set is flabby, thus the law of
excluded middle follows by Proposition~\ref{prop:taboo}.

@@

\begin{proof}We first cover the case of sets. Let~$K \subseteq I^T$ be a
subterminal. We consider the injectivity diagram
\[ \xymatrix{
  T' \ar@{^{(}->}[r]\ar[d] & T \ar@{-->}@/^/[ld] \\
  I
} \]
where~$T'$ is the subset~$\{ s \in T \,|\, \text{$K$ is inhabited} \} \subseteq T$ and the
solid vertical map sends~$s \in T'$ to~$g(s)$, where~$g$ is an arbitrary element
of~$K$. This association is well-defined. Since~$I$ is injective, a dotted lift
as indicated exists. If~$K$ is inhabited, this lift is an element of~$K$.

The same kind of argument applies to the case of modules. If~$K$ is a
subterminal of~$\Hom_R(T,I)$, we define~$T'$ to be the submodule
$\{ s \in T \,|\, \text{$s = 0$ or $K$ is inhabited} \}$ and consider the
analogous injectivity diagram, where the solid vertical map~$f : T' \to I$ is now
defined by cases: Let~$s \in T'$. If~$s = 0$, then we set~$f(s) = 0$; if~$K$ is
inhabited, then we set~$f(s) \defeq g(s)$, where~$g$ is an arbitrary element
of~$K$. This association is again well-defined, and a dotted lift yields the
desired element of~$\Hom_R(T,I)$.
\end{proof}

@@

\begin{proof}Set~$X \defeq \{ \bar{f} \in I^B \,|\, \bar{f} \circ i =
f \}$. Let~$K \subseteq X$ be a subsingleton. We consider the injectivity diagram
\[ \xymatrix{
  i[A] \cup B' \ar@{^{(}->}[r]\ar[d]_g & B \ar@{-->}@/^/[ld] \\
  I
} \]
where~$B'$ is the set~$\{ s \in B \,|\, \text{$K$ is inhabited} \}$ and the solid
vertical arrow~$g$ is defined in the following way: Let~$s \in i[A] \cup B'$.
If~$s \in i[A]$, then we set~$g(s) \defeq f(a)$, where~$a \in A$ is an element such
that~$s = i(a)$. If~$s \in B'$, then we set~$g(s) \defeq \bar{f}(s)$,
where~$\bar{f}$ is any element of~$K$. These prescriptions determine a well-defined
map.

@@

\begin{proof}The proof is similar to the proof of
Proposition~\ref{prop:set-of-extensions-flabby}. We
set~$X \defeq \{ \bar{f} \in \Hom_R(B,I) \,|\, \bar{f} \circ i =
f \}$. Let~$K \subseteq X$ be a subsingleton. We consider the injectivity diagram
\[ \xymatrix{
  i[A] + B' \ar@{^{(}->}[r]\ar[d]_g & B \ar@{-->}@/^/[ld] \\
  I
} \]
where~$B'$ is the submodule~$\{ s \in B \,|\, \text{$s = 0$ or $K$ is
inhabited} \} \subseteq B$ and the solid vertical arrow~$g$ is defined in the following
way: Let~$s \in i[A] + B'$. Then~$s = i(a) + t$ for an element~$a \in A$ and an
element~$t \in B'$. Since~$t \in B'$, $t = 0$ or~$K$ is inhabited. If~$t = 0$,
we set~$g(s) \defeq f(a)$. If~$K$ is inhabited, we set~$g(s) \defeq f(a) +
\bar{f}(s)$, where~$\bar{f}$ is any element of~$K$. These prescriptions
determine a well-defined map.

@@

\begin{proof}Let~$X \defeq \{ u \in M \,|\, p(u) = s
\}$. Let~$K \subseteq X$ be a subterminal. Since~$p$ is surjective, there is an
element~$u_0 \in X$. The translated set~$K - u_0 \subseteq M$ is still a
subterminal, and its preimage under~$i$ is as well. Since~$M'$ is flabby, there
is an element~$v \in M'$ such that~$i^{-1}[K - u_0] \subseteq \{v\}$. We verify
that~$K \subseteq \{u_0 + i(v)\}$.

@@

Noticeably missing here is XXX ...

@@

Instead of referencing arbitrary stages~$A \in \E$, one can also just reference
the generic stage: Let~$\P_{\leq1}(X)$ denote the \emph{object of subterminals}
of~$X$; this object is a certain suboject of~$\P(X) = [X,\Omega_\E]_\E$, the
powerobject of~$X$. The subobject~$K_0$ of~$X \times \P_{\leq1}(X)$ classified by the
evaluation morphism~$X \times \P_{\leq1}(X) \to X \times \P(X) \to \Omega_\E$
is the \emph{generic subterminal} of~$X$. The object~$X$ is flabby if and only
if there exists an epimorphism~$B \to \P_{\leq1}(X)$ and a morphism~$B \to X$
such that the following diagram commutes.
\[ \xymatrix{
  K_0 \times_{\P_{\leq1}(X)} B \ar@{^{(}->}[r]\ar[d] & \P_{\leq1}(X) \ar@{-->}@/^/[ld] \\
  X
} \]

@@

\begin{proof}This holds for any property which can be defined
in the stack semantics~\cite[Lemma~7.3]{shulman:stack-semantics}.
\end{proof}

@@

\begin{prop}\label{prop:global-elements}
Let~$X$ be a flabby object in an elementary topos~$\E$. If~$\E$ is localic and
Zorn's lemma is available, then~$X$ possesses a global element (a morphism~$1 \to X$).
\end{prop}

\begin{proof}This is a restatement of the discussion following
Definition~\ref{defn:flabby-sheaf}.
\end{proof}

@@

\begin{proof}Let~$X \in \F$ be a flabby object.
Let~$k : K \to A$ be a monomorphism in~$\E$ and let~$x : K \to f_*(X)$ be an
arbitary morphism. Without loss of generality, we may assume that~$A$ is the
terminal object~$1$ of~$\E$. Then~$f^*(k) : f^*(K) \to 1$ is a monomorphism
in~$\F$ and the adjoint transpose~$x^t : f^*(K) \to X$
is a morphism in~$\F$. Since~$X$ is flabby, there is an epimorphism~$B \to 1$
in~$\F$ and a morphism~$y : B \to X$ such that the morphism~$f^*(K) \times B
\to X$ factors over~$y$. Hence~$x$ factors over~$f_*(y) : f_*(B) \to f_*(X)$.
We conclude because the morphism~$f_*(B) \to f_*(1)$ is an epimorphism by
assumption.
\end{proof}

The assumption on~$f_*$ of Proposition~\ref{prop:pushforward-of-flabby-objects}
is for instance satisfied if~$f$ is a local geometric morphism or if~$f$ is
induced by an open continuous injection between topological spaces. XXX

@@

\begin{prop}\label{prop:notions-of-internal-injectivity}
Let~$\E$ be an elementary topos. Then the following statements about an
object~$I \in \E$ are equivalent.
% Take care if numbering changes (is referenced below).
\begin{enumerate}
\item[(1)] $I$ is internally injective.
\item[(1')] For any morphism $p : A \to 1$ in $\E$, the object $p^*(I)$ has property~(1)
as an object of $\E/A$.
\item[(2)] The functor~$[\cdot, I] : \E^\op \to \E$ maps monomorphisms in $\E$
to morphisms for which any global element of the target locally (after change of
base along an epimorphism) possesses a preimage.
\item[(2')] For any morphism $p : A \to 1$ in $\E$, the object $p^*(I)$ has property~(2)
as an object of $\E/A$.
\item[(3)] The statement~``$I$ is an injective set'' holds in the stack
semantics of~$\E$.
\end{enumerate}
\end{prop}

\begin{proof}
The implications (1)~$\Rightarrow$~(2), (1')~$\Rightarrow$~(2'),
(1')~$\Rightarrow$~(1) and (2')~$\Rightarrow$~(2) are trivial.

The equivalence (1')~$\Leftrightarrow$~(3) follows directly from the
interpretation rules of the stack semantics.

The implication (2)~$\Rightarrow$~(2') employs the
extra left adjoint $p_! : \E/A \to \E$ of $p^* : \E
\to \E/A$~(which maps an object~$(X \to A)$ to~$X$), as in the usual proof that
injective sheaves remain injective when
restricted to smaller open subsets: We have that $p_* \circ [\cdot, p^*(I)]_{\E/A}
\cong [\cdot, I]_\E \circ p_!$, the functor $p_!$ preserves monomorphisms, and one
can check that $p_*$ reflects the property that global elements locally possess
preimages. Details are in~\cite[Thm.~1.1]{harting}.\footnote{Harting formulates
the statement for abelian group objects, and has to assume that~$\E$ contains a
natural numbers object to ensure the existence of an abelian version of~$p_!$.}

The implication (2')~$\Rightarrow$~(1') follows by performing an extra change of
base, since any non-global element becomes a global element after a suitable
change of base.
\end{proof}

Let~$R$ be a ring in~$\E$. Then the analogue of
Proposition~\ref{prop:notions-of-internal-injectivity} holds for~$R$-modules
in~$\E$, if~$\E$ is assumed to have a natural numbers object. The extra
assumption is needed in order to construct the left adjoint~$p_! :
\Mod_{\E/A}(R \times A) \to \Mod_\E(R)$. Phrased in the internal language, this
adjoint maps a family~$(M_a)_{a \in A}$ of~$R$-modules to the direct
sum~$\bigoplus_{a \in A} M_a$. Details on this construction, phrased in the
language of sets but interpretable in the internal language, can for instance
be found in~\cite[page~54]{mines-richman-ruitenburg:constructive-algebra}.

@@

For the ``if'' direction, let~$I$ be an internally
injective object. Let~$i : A \to B$ be a monomorphism in~$\E$ and let~$f :
A \to I$ be an arbitrary morphism. We want to show that there exists an
extension $B \to I$ of~$f$ along~$i$. To this end, we consider the object of
such extensions, defined by the internal expression
\[ F \defeq \{ \bar{f} \in [B,I] \,|\, \bar{f} \circ i = f \}. \]
Global elements of~$F$ are extensions of the kind we are looking for.
By Proposition~\ref{prop:set-of-extensions-flabby}, this object is flabby.
By Proposition~\ref{prop:global-elements}, it has a global element.
\end{proof}

The analogue of Theorem~\ref{thm:injectivity-external-internal} for modules
holds as well, if~$\E$ is assumed to have a natural numbers object. The proof
carries over word for word, only referencing
Proposition~\ref{prop:set-of-extensions-flabby} instead of
Proposition~\ref{prop:set-of-linear-extensions-flabby}.
It seems that Harting was not aware of this, even though she did show that
injectivity of sheaves of modules over topological spaces is a local
notion~\cite{XXX}, as she
(mistakingly) states in~\cite[page~233]{harting:remark} that ``the notions of
injectivity and internal injectivity do not coincide'' for modules.
% XXX "mistakenly"? comparison with Harting's proof (Barr's lemma...)

@@

\begin{scholium}\label{scholium:properties-of-flabby-objects}
For any elementary topos~$\E$:
\begin{enumerate}
\item Any object can be embedded into an internally injective object.
\item (If~$\E$ has a natural numbers object.) The underlying unstructured
object of an internally injective module is internally injective.
\item Any internally injective object is flabby.
% Take care if numbering changes (referenced below).
\item Any object can be embedded into a flabby object.
\item The terminal object is flabby. The product of flabby objects if flabby.
\item Let~$I$ be an internally injective object. Let~$T$ be an arbitrary
object. Then~$[T,I]$ is a flabby object.
\item (If~$\E$ has a natural numbers object.) Let~$I$ be an internally
injective~$R$-module. Let~$T$ be an arbitrary~$R$-module. Then~$[T,I]_R$, the
subobject of the internal Hom consisting only of the linear maps, is a flabby
object.
\item Let~$0 \to M' \to M \to M'' \to 0$ be a short exact sequence
of~$R$-modules in~$\E$. If~$M'$ and~$M''$ are flabby objects, so is~$M$.
\end{enumerate}
\end{scholium}

@@

\begin{proof}We only have to verify exactness at~$\Gamma(M'')$, so let~$s \in
\Gamma(M')$. Interpreting Proposition~\ref{prop:set-of-preimages-flabby}
in~$\E$, we see that the object of preimages of~$s$ is flabby. Since~$\E$ is
localic, this object is a flabby sheaf; since Zorn's lemma is available, it
possesses a global element. Such an element is the desired preimage of~$s$
in~$\Gamma(M)$.\end{proof}

@@

\begin{rem}A direct generalization of the traditional notion of a flabby sheaf, as
opposed to our reimagining in Definition~\ref{defn:flabby-sheaf}, to
elementary toposes is the following. An object~$X$ of an elementary topos~$\E$
is \emph{strongly flabby} if and only if, for any monomorphism~$K \to 1$
in~$\E$, any morphism~$K \to X$ lifts to a morphism~$1 \to X$.

@@

XXX
% EXT, TOR
% Higher direct images

@@

The main obstacle preventing Grothendieck's theory of derived functors to be interpreted
constructively is its reliance on injective resolutions. It is known that in
the absence of the axiom of choice, much less in a purely intuitionistic
context, there might not be any nontrivial injective abelian
group~\cite{blass:inj-proj-axc}.

In principle, this problem could be remedied by employing flabby resolutions
instead of injective ones. There are, however, two problems with this
suggestion. Firstly, we needed Zorn's lemma to show that flabby sheaves are
acyclic for the global sections functor (Proposition~XXX). This problem might
be mitigated by relying on the substitute property XXX.
The more serious problem is that it is an open question whether
one can show, purely intuitionistically, that any sheaf of modules embeds into
a flabby sheaf of modules. The following is known about this problem:

\begin{enumerate}
\item There is a purely intuitionistic proof that any sheaf of sets embeds into
a flabby sheaf of sets
(Scholium~\ref{scholium:properties-of-flabby-objects}(4)).

\item The existence of enough flabby modules, and even the existence of enough
injective modules, is \emph{not} a constructive taboo, that is, these statements do not
entail a classical principle like the law of excluded middle or the principle
of omniscience. This is because assuming the axiom of choice, any
Grothendieck topos has enough injective (and therefore flabby) modules.
